---
title: "UK Smoking Data"
author: "Lokesh Surendra Jain"
date: '2023-02-21'
output:
  pdf_document: default
  html_document:
    df_print: paged
always_allow_html: true
---

In this exercise we will study UK Smoking Data (`smoking.R`, `smoking.rda` or `smoking.csv`):

**Description**

Survey data on smoking habits from the UK. The data set can be used for analyzing the demographic characteristics of smokers and types of tobacco consumed.

**Format**

A data frame with 1691 observations on the following 12 variables.

`gender` - Gender with levels Female and Male.

`age` - Age.

`marital_status` - Marital status with levels Divorced, Married, Separated, Single and Widowed.

`highest_qualification` - Highest education level with levels A Levels, Degree, GCSE/CSE, GCSE/O Level, Higher/Sub Degree, No Qualification, ONC/BTEC and Other/Sub Degree

`nationality` - Nationality with levels British, English, Irish, Scottish, Welsh, Other, Refused and Unknown.

`ethnicity` - Ethnicity with levels Asian, Black, Chinese, Mixed, White and Refused Unknown.

`gross_income` - Gross income with levels Under 2,600, 2,600 to 5,200, 5,200 to 10,400, 10,400 to 15,600, 15,600 to 20,800, 20,800 to 28,600, 28,600 to 36,400, Above 36,400, Refused and Unknown.

`region` - Region with levels London, Midlands & East Anglia, Scotland, South East, South West, The North and Wales

`smoke` - Smoking status with levels No and Yes

`amt_weekends` - Number of cigarettes smoked per day on weekends.

`amt_weekdays` - Number of cigarettes smoked per day on weekdays.

`type` - Type of cigarettes smoked with levels Packets, Hand-Rolled, Both/Mainly Packets and Both/Mainly Hand-Rolled

Source National STEM Centre, Large Datasets from stats4schools, <https://www.stem.org.uk/resources/elibrary/resource/28452/large-datasets-stats4schools>.

Obtained from <https://www.openintro.org/data/index.php?data=smoking>

## Read and Clean the Data 

```{r setup, results="hide", warning=F, message=F}
# load libraries

#install.packages("factoextra")
#install.packages("FactoMineR")
#install.packages("missMDA")

if (!require("ggbiplot")) {
  if (!require("devtools")) install.packages("devtools")
  devtools::install_github("vqv/ggbiplot")
}

library(dplyr)
library(FactoMineR)
library(factoextra)

library(ggplot2)
library(ggbiplot)
library(fastDummies)

library(plotly)
setwd("C:/Users/lokes/Downloads")

```

```{r}
# Load data

source("smoking.R")

```

Take a look into data

```{r}
# place holder
smoking <- smoking
smoking

```

There are many fields there so for this exercise lets only concentrate on smoke, gender, age, marital_status, highest_qualification and gross_income.

Create new data.frame with only these columns.

```{r}
# place holder

smoking_subset <- smoking[, c( "smoke", "gender", "age", "marital_status", "highest_qualification", "gross_income")]
head(smoking_subset)
```

```{r}
# place holder

smoking_subset <- na.omit(smoking_subset)

new_nrows <- nrow(smoking_subset)
head(new_nrows)


```

```{r}
# place holder

smoking_subset$gender <- ifelse(smoking_subset$gender == "Male", 1, 0) # Female = 0, Male = 1
smoking_subset$smoke <- ifelse(smoking_subset$smoke == "No", 0, 1) # No = 0, Yes = 1

head(smoking_subset$gender)
head(smoking_subset$smoke)

smoking_subset$gross_income <- as.integer(as.factor(smoking_subset$gross_income))
smoking_subset$highest_qualification <- as.integer(as.factor(smoking_subset$highest_qualification))

head(smoking_subset$gross_income)
head(smoking_subset$highest_qualification)

smoking_subset <- smoking_subset %>%
  mutate(
    is_married = if_else(marital_status == "Married", 1, 0),
    is_divorced = if_else(marital_status == "Divorced", 1, 0),
    is_single = if_else(marital_status == "Single", 1, 0),
    is_widowed = if_else(marital_status == "Widowed", 1, 0)
  )%>%
select(-marital_status)

glimpse(smoking_subset)


#any(is.na(smoking_subset))  # check any missing value
 
```

```{r}
# place holder

# smoking_subset <- subset(smoking_subset, select = -c(smoke))

pca_result <- prcomp(smoking_subset[c(-1)], scale. = TRUE)

summary(pca_result)

 
```

```{r}

plot(pca_result, type = "line")
# place holder
```

```{r echo=TRUE}

ggbiplot(pca_result,scale=0,groups=as.factor(smoking_subset$smoke))
# place holder
```

```{r}

smoking_subset$smoke <- factor(
  smoking_subset$smoke,
  levels = c(1, 2, 3),
  labels = c("never", "former", "current")
)

pca_result <- prcomp(smoking_subset[, -1], scale. = TRUE)

library(ggbiplot)
ggbiplot(
  pca_result,
  obs.scale = 1,
  var.scale = 1,
  groups = smoking_subset$smoke,
  ellipse = TRUE,
  circle = TRUE
)



# place holder
```

-   Convert a columns to proper format

```{r}

setwd("C:/Users/lokes/Downloads")
ds_salary <- read.csv("ds_salaries.csv")
ds_salary

ds_salary_subset <- ds_salary[, c( "experience_level", "employment_type", "job_title", "company_size",'work_year')]
head(ds_salary_subset)


ds_salary_subset <- na.omit(ds_salary_subset)

ds_salary_subset$employment_type <- ifelse(ds_salary_subset$employment_type == "FT", 1, 0)

ds_salary_subset$company_size <- as.integer(as.factor(ds_salary_subset$company_size))


ds_salary_subset$experience_level <- as.integer(as.factor(ds_salary_subset$experience_level))

ds_salary_subset$job_title <- as.integer(as.factor(ds_salary_subset$job_title))

```

-   Perform PCA

    ```{r}

    head(sapply(ds_salary_subset, class))

    pca_result_ds_salary <- prcomp(ds_salary_subset, scale. = TRUE)

    summary(pca_result_ds_salary)

    ```

-   Make a skree plot

    ```{r}

    plot(pca_result_ds_salary, type = "line")
    ```

-   Make a biplot

    ```{r}

    ggbiplot(pca_result_ds_salary,scale=0,groups=as.factor(ds_salary_subset$company_size))

    ```
